Towards Semantic Patches

### 2010-06-18 papers/MIPS2010/abstract.tex
* Some quickly written paragraphs.
git-svn-id: http://gasp.dev.regis-gianas.com/svn/trunk@102 403e655e-980a-47bd-8c60-1ee5348bb6b7

@@

>>>>

The kernel language described in the previous section is clearly not
adapted to human interaction, for several reasons~: 
\begin{itemize}

\item Every single object is flattened, which makes unobvious the
  structure of a program or a proof.

\item A term of this language denotes the entire history of a
  development. Most of the time, the user is more interested in a
  instantaneous snapshot of his object of study.

\item Even though we were able to provide a snapshot of a particular
  state of the development, the repository represents this state 
  using proof terms, which contain too many syntactic details from the
  user point of view and are not suitable for edition.

\end{itemize}

@@

With the previous remarks in mind, we designed a generic framework of semantic
patches that can be described using three layers which are summarized in 
Figure~\ref{fig:architecture}.

The most internal layer focused on the history of metatheory
objects. It is composed of a type-checker for the kernel language. The
purpose of this layer is to maintain a repository~$R$ and the
invariant that it is well-typed. Through the layer interface, a client
can ask for the \textit{integration} $\theta(R)$ of a patch $\Delta$
to obtain an extended version $R'$ of the repository $R$.

The middle layer is a mechanized formalization of a metatheory. This
module provides a syntax for the objects of the metatheory (formulae,
judgments, \ldots) as well as the admissible rules (that we also call
metatheorems).  An \textit{internalization} function $\phi$ translates
the constructors of these objects and the specification of the
admissible rules as terms of the kernel language. As a consequence,
every object~$t$ of the theory can be injected into the
repository~$R$. An \textit{externalization} function $\psi$ is
implementing the inverse operation. Given a modified version $t'$ of
an object~$t$, we are able to compute the patch $\Delta$ using
the \textit{differienciation} function $\nabla(t)$.

The external layer offers views on the objects of the theory. A
view~$w$ is a human readable, and editable, partial representation of
an object~$t$. A view is obtained from an object through a
projection~$\pi$.  The user transforms the view $w$ into a view $w'$
which must be $\textit{put back}$ into the metatheory as a modified
object $t'$ using a $\textit{putback}$ function $\sigma(t, w)$.
